---
title: "Looking into my mouth microbiome"
author: "Richard Sprague @sprague"
date: "February 25, 2015"
output: html_document
---
The gut biome is interesting enough, but bacteria colonize just about every part of the body, so recently I've been studying [my uBiome mouth test results](http://blog.richardsprague.com/2014/10/whats-in-my-microbiome.html). The simple [GitHub RuBiome utilities](https://github.com/richardsprague/uBiome) I use for analyzing my gut will work for that too, so here's a short example of how I did it:

```{r, echo=FALSE}
.myuBiomePath = "OneDrive/Projects/"
# set the root path for either Mac or Windows
if(Sys.info()["sysname"]=="Darwin") kRootPath=paste0("~/",.myuBiomePath) else kRootPath=paste0(paste0("c:/Users/",Sys.info()["login"]),.myuBiomePath)

setwd(paste0(kRootPath,"uBiome"))
#setwd("~/OneDrive/Projects/uBiome")
# load the uBiome utilities
source("uBiome.R")

# Now load all the sprague data
setwd("sprague data")
topN <- as.integer(5) # a global integer to pick the number of results I'll display

junMouth<-read.csv("junMouth.csv")
OctMouth<-read.csv("OctMouth.csv")

```

First I [loaded my uBiome data](http://blog.richardsprague.com/2015/01/how-to-analyze-ubiome-sample-in-excel.html) into two variables, one for each sample: June 2014 (```junMouth```) and the other from October 2014 (```OctMouth```), after [a visit to my dentist](http://blog.richardsprague.com/2014/10/why-review-dentists.html).

Let's see which species are new in the later (October) sample:


```{r}
octToJunChange <- uBiome_sample_unique(OctMouth,junMouth)
```
```{r,echo=FALSE}
names(octToJunChange)[1]<-"count"
head(octToJunChange,topN)
```

One of those species, *Streptococcus sanguinis* looks interesting. [Wikipedia](https://en.wikipedia.org/wiki/Streptococcus_sanguinis) says this:

>*S. sanguinis* is a normal inhabitant of the healthy human mouth where it is particularly found in dental plaque, where it modifies the environment to make it less hospitable for other strains of Streptococcus that cause cavities, such as Streptococcus mutans.

No cavities? Nice! More good news: this quick check confirms that I don't have any *S. mutans*:
```{r}
OctMouth[grepl("Streptococcus",OctMouth$tax_name),]$tax_name
```

Then I look at the species that disappeared (went extinct) between the two samples:
```{r}
junToOctChange <- uBiome_sample_unique(junMouth,OctMouth)
```
```{r,echo=FALSE}
names(junToOctChange)[1]<-"count"
head(junToOctChange[order(junToOctChange$count,decreasing=TRUE),],topN)
```
Anything in the genus [*Capnocytophaga*](https://en.wikipedia.org/wiki/Capnocytophaga) is an opportunistic pathogen, so I say good riddance. Usually they're fine, but if your immune system dips they can turn bad.

[Streptococcus rubneri](http://www.ncbi.nlm.nih.gov/pubmed/23749274) was discovered a couple years ago, so little is known about it.

[Prevotella buccae](https://microbewiki.kenyon.edu/index.php/Prevotella) is more interesting. It seems to be implicated in periodontal disease (yuk!) but that genus is involved too in breaking down proteins and carbohydrates.

Hard to say what's really going on.  Meanwhile, here are the biggest changes (increase) since the first sample:
```{r}
junToOctCompare <- uBiome_compare_samples(junMouth,OctMouth)
```
```{r,echo=FALSE}

head(junToOctCompare[order(junToOctCompare$count_change,decreasing=TRUE),],topN)
```
Whoah! That first one, *Streptococcus pseudopneumoniae*, looks nasty! [Wikipedia](https://en.wikipedia.org/wiki/Streptococcus_pseudopneumoniae) says it may cause pneumonia, though a recent medical journal says more hopefully that it ["treads the fine line between commensal and pathogen"](http://www.cmnewsletter.com/article/S0196-4399(14)00027-0/abstract?cc=y)

which is a scientific gobbleygook way of saying nobody has a clue. All the more reason to keep testing, submitting, and getting more data.  I just sent two more kits to uBiome, and will let you know more as soon as I get back the results.
