---
title: "How to use RuBiome Utilities"
author: "Richard Sprague @sprague"
date: "February 4, 2015"
output: html_document
---

##Important Note##

You won't learn much from this session. It needs much cleaning up to be useful. Although it does actually work, I'm keeping it around for my own use later. You're better off using the other documentation on this site.

This is an interactive session that shows how I work with the utilities.
You'll first need to load everything.

```{r}

setwd("../Data") # set your working directory to the place where you keep your data.

source("../sources/uBiomeCompare.R")
may<-read.csv("sprague-uBiomeMay2014.csv")
jun<-read.csv("sprague-uBiomeJun2014.csv")
oct<-read.csv("sprague-uBiomeOct2014.csv")
```

My data for June has weird headings for some reason. Maybe uBiome changed the way they write JSON for that particular file. Anyway, to make things consistent I need to clean up the headers.

```{r}

junNames<-names(jun)
junNamesS<-strsplit(junNames,"\\.")
names(junNamesS)<-NULL
junNamesT<-sapply(junNamesS,function(x){x[2]})
names(jun)<-junNamesT


names(jun)
names(oct)

```

Next, look just at the species in my samples:

```{r}
maySpecies <-may[may$tax_rank=="species",]
octSpecies <-oct[oct$tax_rank=="species",]
```
...and sort them alphabetically
```{r}
#sort
msA<-maySpecies[order(maySpecies$tax_name),] #May species table alphabetized 
osA<-octSpecies[order(octSpecies$tax_name),] #Oct species table alphabetized
```

Find all the May species that are still in the October sample:
```{r}
# maySpecies %in% octSpecies
mayOctMatch<-match(maySpecies$tax_name,octSpecies$tax_name)
ms<-which(maySpecies$tax_name %in% octSpecies$tax_name)
os<-which(octSpecies$tax_name %in% maySpecies$tax_name)
maySpeciesstillInOct <-maySpecies$tax_name[ms] # names of all May species still found in Oct sample
mayt<-maySpecies[ms,] # full table of all May species still found in Oct
octt<-octSpecies[os,]
#important: note that rownnames(octt) maintains references to the original row names from oct
# sapply(rownames(mayt),as.integer) #to convert to a vector, instead of char
maytA<-mayt[order(mayt$tax_name),] #alphabetized version of mayt
octtA<-octt[order(octt$tax_name),] #alphabetized version of octt
```

and compute the differences between May and October:

```{r}

species_change_May_Oct <-data.frame(maytA$tax_name,octtA$count_norm - maytA$count_norm)
```

Finally, write everything to another CSV file that you can read into Excel:

```{r}
#write.csv(species_change_May_Oct,file="Sprague-uBiome change May-Oct.csv")
```



