---
title: "Jungle microbiome details"
author: "Richard Sprague @sprague"
date: "April 22, 2015"
output: html_document
---



```{r, echo=FALSE}
setwd("../Data")
source("../sources/ubiomeCompare.R")
jan<-read.csv("sprague-ubiomejan2015x.csv")

feb<-read.csv("sprague-ubiomeFeb2015.csv")
# compare my trip to Belize
# positive numbers mean more because i went to Belize
janFeb<-uBiome_compare_samples(jan,feb)
janFebGenus<-uBiome_compare_samples(jan,feb,"genus")
#write.csv(janFebGenus,"belizeTripGenus.csv")

newAfterBelizeGenus<-uBiome_sample_unique(feb,jan,"genus")  # found in feb, but not jan
extinctAfterBelize<-uBiome_sample_unique(jan,feb,"genus")

```

Using my [free public uBiome analysis tools](https://github.com/richardsprague/uBiome), I see that I picked up the following genuses while in the jungle:

```{r}

head(newAfterBelizeGenus)
```

and a few that went extinct:
```{r}
head(extinctAfterBelize)
```

What do these bacteria do? As always, who knows? The science is so new that when I look up most of these genuses online and elsewhere, I find almost nothing relevant. All the more reason I hope you'll submit a sample to uBiome the next time *you* travel in the jungle, so we can compare and learn together.

